---
title: "hw4"
author: 'Thomas Le, ID: 91308173 and Armand Nasseri, ID: 912679383'
date: "June 9, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1

## a)

Supposing process begins at state 1 at T=1. Probability it will be at state 0 at t=3 is the sum of:

```
1 -> 0 -> 0 = .7 * .4 = 0.28
1 -> 1 -> 0 = .3 * .7 = 0.21

Then, the answer is 0.28 + 0.21 = 0.49
```

## b)

Solving for steady-state distribution solving for $\mu P = \mu$ using eigenvalues

Source: https://nicolewhite.github.io/2014/06/10/steady-state-transition-matrix.html

```{r, echo=TRUE}
# markov chain table setup
p = matrix(c(0.4, 0.7,
             0.6, 0.3),
           nrow = 2,
           ncol = 2)

states = c(0, 1)
colnames(p) = states
rownames(p) = states
p

# solving using eigen values
e = eigen(t(p))
first = Re(e$vectors[, 1])

# normalize
mu = first / sum(first)

mu
```

# 2

Given mean a Poisson distribution with mean of 4.2 bugs per 1,000 lines of code.
Searching for probability of more than 40 bugs in 20 sections of code, each 1,000
lines long, where sections are independent.

For a Poisson distribution, E[X] = $\lambda$. Given that the mean is 4.2, then $\lambda$ = 4.2

For a set of 20 sections, then the mean is now `E[X] = 4.2 * 20 = 84`. 

Similarily, `Var[X] = 4.2 * 20 = 84`.

Then, for a set of i.i.d random variables, its distribution is approximate normal, $N(84, \sqrt{84})$.

Then, P(X > 40) = P(z > $\frac{40-84}{\sqrt{84}}$) = P(z > -4.80). Though -4.8 is
not on the z table, the z value is very far to the left of the distribution so
the p value should be approximate 0.0001. Then the probability is $1-0.0001 = 0.9999$.

The r function to find the z value would be $pnorm(40, 84, \sqrt{84})$. 

So the answer using R would be $1 - pnorm(40, 84, \sqrt{84}) = 1 - 7.933e \text{-} 07 = 0.9999$

# 3

Given an exponential distribution with mean = 15, then because for an exponential distribution, 
$E[X] = \frac{1}{\lambda}$, then $\lambda = \frac{1}{15}$.


```{r, echo=TRUE}
# generate 10 random numbers
lambda <- 1/15
weights <- rexp(10, lambda)
cat("Randomized weights are:") 
weights

# for a 93% CI, due to symmetry there is 3.5% on left and right side, so we search for Z val at 0.965
z <- qnorm(0.965)

xbar <- mean(weights)

s2 <- mean(weights^2) - xbar^2
s <- sqrt(s2)
radius <- z * s/sqrt(10)
cat("Approximate confidence interval is", xbar-radius, "to", xbar+radius, "\n")

```


# 4

To test that the length of the two fish (salmon S, tuna T) have no difference against $\alpha = 0.05$, let:

$$ H_0: \overline{S} - \overline{T} = 0$$
$$ H_A: \overline{S} - \overline{T} \neq 0$$


```{r, echo=TRUE}
# salmon s
sbar <- 14
sstd <- 2
ssize <- 20

# tuna t
tbar <- 15
tstd <- 3
tsize <- 15

# standard error
se <- sqrt( ((sstd^2)/ssize) + ((tstd^2)/tsize))

Z <- (sbar-tbar-0)/se

# p value
pval <- 2*pnorm(Z)
pval

```

The P-value for the test statistic is 0.2635525, which is not less than the alpha value of 0.05.

Thus, we fail to reject the null hypothesis at the 0.05 significance level.

In context, we fail to reject the hypothesis that there is no difference in the length of salmon vs tuna in the sacramento river.





