---
title: "HW3"
author: "Thomas Le, ID: 913091973 ; Armand Nasseri ID: 912679383"
date: "May 30, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We are taking one of our free late days.

# A

```{r, echo=TRUE}
sim1 <- function(nreps, p=0.2, k=3) {
  # get rand distrubtion of nreps size
  N <- rgeom(nreps, p)
  
  # add 1 since we have k = 1, 2 ...
  N <- N + 1
  
  # result is N where N <= k
  res <- N[N <= k]
  return(mean(res))
}
```

# B

Cov(D, N) = E[DN] - (E[D] * E[N])

$E[DN] = \sum_{i=1}^{10} i(11-i)(1-0.15)^{i-1}(0.15) + \sum_{i=11}^{\infty} i(i-11)0.85^{i-1}0.15$

$E[D] = \sum_{i=1}^{10} (11-i)(1-0.15)^{i-1}(0.15) + \sum_{i=11}^{\infty} (i-11)0.85^{i-1}0.15$

$E[N] = \frac{1}{p}$ 

Then, $Cov(D, N) = \sum_{i=1}^{10} i(11-i)(1-0.15)^{i-1}(0.15) + \sum_{i=11}^{\infty} i(i-11)0.85^{i-1}0.15 - [\sum_{i=1}^{10} (11-i)(1-0.15)^{i-1}(0.15) + \sum_{i=11}^{\infty} (i-11)0.85^{i-1}0.15 * \frac{1}{p}]$

Plugging in p = 0.15 into a calculator...=>

(17.14228235171777 + 26.031171240606657) - ((5.448954957930762 + 1.115621624597426) * (6.666667)) = -0.5903925


# C

```{r, echo=TRUE}
library(plyr)

# get data, skip header row
original_data <- read.table("./dnc-corecipient/out.dnc-corecipient", skip=1)

# rename cols 1 and 2
colnames(original_data) <- c("id1", "id2", "nummsgs")

# select rows where id1 < id2 to remove duplicate data, keeping only cols 1 and 2
dnc <- original_data[original_data$id1 < original_data$id2, c(1,2)]

# create empty vector of 0s; get degrees by counting num occurences of each val
degrees <- rep(0, nrow(dnc))
for (i in 1:nrow(dnc)) {
  degrees[dnc[i, "id1"]] <- degrees[dnc[i, "id1"]] + 1
  degrees[dnc[i, "id2"]] <- degrees[dnc[i, "id2"]] + 1
}

# get max degree, to know what i goes up to
max_degree <- max(degrees)
  
# mi is count of recipients having degree i
mi <- rep(0, max_degree)
for (i in 1:nrow(dnc)) {
  # for non zero degrees
  if (degrees[i] != 0) {
    # increment its count
    mi[degrees[i]] <- mi[degrees[i]] + 1
  }
}

# since calling log on 0 returns -Inf, we replace 0s with NA so plot will ignore the NA vals
for (i in 1:length(mi)) {
  if (mi[i] == 0)
    mi[i] <- NA
}

# i goes from 1 to max degree
i <- c(1:max_degree)

# apply log to i and mi
log_i <- log(i)
log_mi <- log(mi)

# apply linear model function and plot
lm(log_i ~ log_mi)
plot(log_i, log_mi)
abline(lm(log_i ~ log_mi))

summary(lm(log_i ~ log_mi))
```

An estimate for gamma was found by applying a linear model function onto the
logarithmic values of m_i and i, where m_i denotes the count of recipients
having degree i in the data, i = 1, 2, 3 ... . 

Gamma is estimated from the linear model to be about -0.8616, the slope from the
linear fit. Interpreting that in context of the data means that as the degree i
grows larger (i.e. number of times a recipient is inolved in a message), then
m_i (count of unique recipients being involved in i messages) grows smaller.
In simpler terms, it is a small amount of people that are involved in the most messages, 
which matches the description of a power law.

Viewing the summary of the linear model, we see that the p-value of 2.2e-16 is
less than 0.05. Thus, the model is statistically significant and that the data
fits the power law distribution well enough.

